Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence System

### 2020-02-26 Revisions/Beauchesne-eDriversMS.Rmd
include final revisions to the manuscript before sending to co-authors

---
title: Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence System
author: David Beauchesne^1,2,\*^, Rémi M. Daigle^2^, Steve Vissault^3^, Dominique Gravel^3^, Andréane Bastien^4^, Simon Bélanger^5^, Pascal Bernatchez^5^, Marjolaine Blais ^6^, Hugo Bourdages^6^, Clément Chion^7^, Peter S. Galbraith^6^, Benjamin Halpern^8,9^, Camille Lavoie^2^, Christopher W. McKindsey^6^, Alfonso Mucci^10^, Simon Pineault^11^, Michel Starr^6^, Anne-Sophie Ste-Marie^4^, Philippe Archambault^2^
fontsize: 12pt
font-family: Times New Roman
output:
  pdf_document:
    toc: false
    number_sections: true
    keep_tex: true
  word_document:
    toc: false
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{listings}
   - \usepackage[svgnames]{xcolor}
   # - \usepackage{titling}
   # - \pretitle{\begin{flushleft}}
   # - \posttitle{\end{flushleft}}
   - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
   - \usepackage{bbm}
   - \usepackage{float}
   - \usepackage{ulem}
   - \definecolor{blue}{HTML}{20cad5}
relativeurls: true
bibliography: Beauchesne-eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

@@

\textcolor{red}{\sout{While the focus of our paper has been on the description of drivers in the St. Lawrence, drivers are but one of the 4 elements required to properly evaluate environmental impacts. This single knowledge node should be weaved with other, comparable, nodes required for integrated management approaches such as species distribution and marine protected areas. Ultimately, all of these knowledge nodes could be weaved together through social-ecological metanetworks (Dee et al. 2017).}}

\textcolor{blue}{A caveat to our initiative is that it focuses on a single element required for fully operational impact assessments. Data that provide knowledge on the exposure of ecosystems to drivers are called stressor-based indicators (Dube and Munkittrick 2001; Dubé 2003; Jones 2016). These indicators efficently identify potential local impacts and can be proactively linked to decision-making, yet assume complete knowledge of drivers and fail to diagnose impacts on valued components or non-additive effects (Dube and Munkittrick 2001; Dubé 2003). In contrast, effect-based indicators are direct measurements of valued components (*e.g.* species abundance and biodiversity) and inherently capture the effects of multiple drivers (Dube and Munkittrick 2001; Dubé 2003). While effect-based indicators are considered superior to stressor-based indicators, they fail to ascribe observed effects to specific drivers (Dube and Munkittrick 2001; Dubé 2003). Stressor-based and effect-based indicators are therefore both required to diagnose causes of ecosystem change (Dube 2003; Jones 2016). As a knowledge node on stressor-based indicators, *eDrivers* should be weaved with other, comparable, nodes describing valued ecosystem components that can be linked to drivers and allow a better understanding of cumulative impacts. Ultimately, multiple knowledge nodes could be weaved together through social-ecological metanetworks analyses (Dee et al. 2016).} <!-- [@dube2001; @dube2003; @jones2016; @dee2017] -->

@@

>>>>

---
title: Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence System
author: David Beauchesne^1,2,\*^, Rémi M. Daigle^2^, Steve Vissault^3^, Dominique Gravel^3^, Andréane Bastien^4^, Simon Bélanger^5^, Pascal Bernatchez^5^, Marjolaine Blais ^6^, Hugo Bourdages^6^, Clément Chion^7^, Peter S. Galbraith^6^, Benjamin Halpern^8,9^, Camille Lavoie^2^, Christopher W. McKindsey^6^, Alfonso Mucci^10^, Simon Pineault^11^, Michel Starr^6^, Anne-Sophie Ste-Marie^4^, Philippe Archambault^2^
fontsize: 12pt
font-family: Times New Roman
output:
  pdf_document:
    toc: false
    number_sections: true
    keep_tex: true
  word_document:
    toc: false
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{listings}
   - \usepackage[svgnames]{xcolor}
   - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
   - \usepackage{bbm}
   - \usepackage{float}
   - \usepackage[normalem]{ulem}
   - \definecolor{blue}{HTML}{20cad5}
relativeurls: true
bibliography: Beauchesne-eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

@@

\textcolor{red}{\sout{While the focus of our paper has been on the description of drivers in the St. Lawrence, drivers are but one of the 4 elements required to properly evaluate environmental impacts. This single knowledge node should be weaved with other, comparable, nodes required for integrated management approaches such as species distribution and marine protected areas. Ultimately, all of these knowledge nodes could be weaved together through social-ecological metanetworks (Dee et al. 2017).}}
\textcolor{blue}{A caveat to our initiative is that it focuses on a single element required for fully operational impact assessments. Data that provide knowledge on the exposure of ecosystems to drivers are called stressor-based indicators (Dube and Munkittrick 2001; Dubé 2003; Jones 2016). These indicators efficently identify potential local impacts and can be proactively linked to decision-making, yet assume complete knowledge of drivers and fail to diagnose impacts on valued components or non-additive effects (Dube and Munkittrick 2001; Dubé 2003). In contrast, effect-based indicators are direct measurements of valued components (\textit{e.g.} species abundance and biodiversity) and inherently capture the effects of multiple drivers (Dube and Munkittrick 2001; Dubé 2003). While effect-based indicators are considered superior to stressor-based indicators, they fail to ascribe observed effects to specific drivers (Dube and Munkittrick 2001; Dubé 2003). Stressor-based and effect-based indicators are therefore both required to diagnose causes of ecosystem change (Dube 2003; Jones 2016). As a knowledge node on stressor-based indicators, \textit{eDrivers} should be weaved with other, comparable, nodes describing valued ecosystem components that can be linked to drivers and allow a better understanding of cumulative impacts. Ultimately, multiple knowledge nodes could be weaved together through social-ecological metanetworks analyses (Dee et al. 2016).}

@@

<!-- !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! -->
TODO: Remove [@dube2001; @dube2003; @jones2016]
<!-- !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! -->
